---
title: "Model Types"
author: "BSB"
date: '2020-03-17'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 

```{r }
setwd("/Users/brandybiggar/Desktop/")
CRAB <- read.csv("egc_working_data_20.csv", header = TRUE)
```

## GlmmTMB

Sometimes glmmTMB, and the families we use with this model require specific data format 
In one case we may want the data to be 0-1 
Sometimes the data needs to be specifically >0 and <1 so we can use the following transformation

Here I create a new variable (abundance, beta.abun) so that I can go back and forth between the different data formats as I try different model components

```{r}
library(glmmTMB)

three$abundance<-three$stand.den/100

y.transf.betareg <- function(y){
  n.obs <- sum(!is.na(y))
  (y * (n.obs - 1) + 0.5) / n.obs
}

three$beta.abun<- y.transf.betareg(three$abundance)
```


# Models by Site

```{r}

new1<-glmmTMB(stand.den~year +(year|site.name),ziformula=~1,data=quinn,family=genpois)
new2<-glmmTMB(stand.den~year +(year|site.name),ziformula=~1,data=two,family=genpois)
new4<-glmmTMB(round(stand.den)~year +(year|site.name),ziformula=~1,data = four,family=genpois)

new3<-glmmTMB(round(beta.abun)~year +(year|site.name),data=three,zi=~beta.abun,family=nbinom2)
summary(new3)

new3<-glmmTMB(log(stand.den+1)~year+ (year|site.name), data=two)


m<-glmmTMB(round(stand.den+1)~year + (1|study.name), data=CRAB, family=nbinom2)
m2<-lmer(round(stand.den) ~ year + (1| study.name), data= CRAB)
qqnorm(residuals(m), main="n")


# don't do nbinom1!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
# nbinom2 has convergence problem 
# don't run compois!
# truncated_compois --- 'stand.den' contains zeros (or values below the allowable range). Zeros are compatible with a trucated distribution only when zero-inflation is added
# genpois has convergence problem
# beta_family has to be between 0-1
# beta_family with beta.abun goes cray

```

