---
title: "Modelling biotic and abiotic through time"
author: "BSB"
date: '2020-03-23'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("/Users/brandybiggar/Desktop/")
CRAB<-read.csv("egc_working_data_21.csv")
```

# Prepare data

Transform standardized abundances (0-100) by rounding (because poisson requires integers).
Divide by 100 because beta is proportions.
Beta transform because beta requires greater than zero and less than 1. 

```{r }

CRAB$round.abun<- round(CRAB$stand.den)
CRAB$prop<- (CRAB$stand.den)/100

y.transf.betareg <- function(y){
  n.obs <- sum(!is.na(y))
  (y * (n.obs - 1) + 0.5) / n.obs
}
CRAB$beta<-y.transf.betareg(CRAB$prop)

```


# Picking a family

I see from this plot that mean is ~= to SD --> Poisson or NB? 
When I look at mean vs SD studies 5, 11, and 12 stand out; keep in mind as potential problem makers.


```{r}
library(dplyr)

a<-CRAB %>% 
  group_by(unique.site.id) %>%
  summarise (mean = mean(stand.den), SD= sd(stand.den))

plot(a$SD, a$mean) 

CRAB2<-filter(CRAB, studyID != "5" & studyID != "11" & studyID != "12")
b<-CRAB2 %>% 
  group_by(unique.site.id) %>%
  summarise (mean = mean(stand.den), SD= sd(stand.den))

plot(b$SD, b$mean) 

```


# Model

Use a loop to model every site.
confint() isn't working so using coef to get sd, will create CIs in next step. 

```{r}
sname=data.frame(unique(CRAB$site.name));rownames(sname)<-sname[,1]
mat<-matrix(0,dim(sname)[1],3)

for (i in 1:dim(sname)[1]) {
  i=3
 dat<-subset(CRAB,site.name==rownames(sname)[i])
  m1<-glm(data=dat, beta~year, family = binomial(link='logit')) 
  mat[i,1:2]<-c(summary(m1)$coefficients[2],summary(m1)$coefficients[4]) # estimate, sd
}
```

# Clean up 

Label columns, add CIs.

```{r}
crabs<-as.data.frame(mat)
colnames(crabs)<- c("Estimate", "sd", "lowCI")
crabs$lowCI <- (crabs$Estimate - crabs$sd)
crabs$upCI <- (crabs$Estimate + crabs$sd)
```


# Explore

Histogram of sd shows that one is nearly 70, this doesn't look good. 
This is studyID 7; wilkinson; 7.61; 61; [-3.5,-3) [54.5,55)

```{r}
hist(crabs$Estimate)
hist(crabs$sd)
hist(crabs$lowCI)
hist(crabs$upCI)
plot(crabs$sd, crabs$Estimate) 

seven<-filter(CRAB, studyID=="7")
weird<-filter(seven, unique.site.id == "7.61")
table(weird$year)
ggplot() + geom_point(data=weird, aes(x=year, y=abun, alpha=1/10))

newdat<-filter(crabs, sd<50)
hist(newdat$sd)
plot(newdat$sd, newdat$Estimate) 
```

Would prefer to remove that site but need to figure out solid justification.


## Temperature

Begin by determining family - gaussian?

```{r}
hist(CRAB$HADiSST.Max_1)
mean(CRAB$HADiSST.Max_1)
var(CRAB$HADiSST.Max_1)

hist(CRAB$HADiSST.Min_1) 
mean(CRAB$HADiSST.Min_1)
var(CRAB$HADiSST.Min_1)

hist(CRAB$NAirT.Min) 
mean(CRAB$NAirT.Min)
var(CRAB$NAirT.Min)
```




Now repeat modelling loop with temperature ~ time


```{r}
sname=data.frame(unique(CRAB$site.name));rownames(sname)<-sname[,1]
mat<-matrix(0,dim(sname)[1],4)

for (i in 1:dim(sname)[1]) {
  i=4
  dat<-subset(CRAB,site.name==rownames(sname)[i])
  m1<-lm(data=dat, NAirT.Max~year)
  mat[i,1:2]<-c(summary(m1)$coefficients[2],summary(m1)$coefficients[4]) # estimate, sd
  mat[i,3]<- mat[i,1] - mat[i,2]
  mat[i,4]<- mat[i,1] + mat[i,2]
}

temps<-as.data.frame(mat)
colnames(temps)<- c("HADMinEst", "HADMinsd", "HADMinlowCI","HADMinupCI") ## HAD Min
temps$site<- rownames(sname)
df2<-temps[,c(5,1,2,3,4)]

temps2<-as.data.frame(mat)
colnames(temps2)<- c("HADMaxEst", "HADMaxsd", "HADMaxlowCI","HADMaxupCI") ## HAD Max

temps3<-as.data.frame(mat)
colnames(temps3)<- c("AirMinEst", "AirMinsd", "AirMinlowCI","AirMinupCI") ## Air Min

temps4<-as.data.frame(mat)
colnames(temps4)<- c("AirMaxEst", "AirMaxsd", "AirMaxlowCI","AirMaxupCI") ## Air Max

tempdf<-cbind(df2, temps2, temps3, temps4)

colnames(crabs)<- c("AbunEst", "Abunsd", "AbunlowCI", "AbunupCI")
slopes.df<-cbind(tempdf, crabs)

setwd("/Users/brandybiggar/Desktop/")
write.csv(slopes.df, "slope.data.march.csv")
```

  m1<-lm(data=dat, HADiSST.Min_1~year)
  m2<-lm(data=dat, HADiSST.Max_1~year)
  m3<-lm(data=dat, NAirT.Min~year)
  m4<-lm(data=dat, NAirT.Max~year)
