---
title: "MS Models May 20"
author: "BSB"
date: '2020-05-10'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("/Users/brandybiggar/Desktop/")
CRAB<-read.csv("egc_working_data_28.csv")
```

## Multi-stage model

# 1. Model abundance through time and temperature through time, separetely for each site. 
  b. Extract the slope of those models. 

Abundance model

```{r eval=FALSE}
sname=data.frame(unique(CRAB$sitename.studyname));rownames(sname)<-sname[,1]
mat<-as.data.frame(matrix(0,dim(sname)[1],21))
colnames(mat)<- c("PoisAbunSlope", "PoisAbunSD","QuasAbunSlope", "QuasAbunSD", "NBAbunSlope", "NBAbunSD", "GLMAbunSlope", "GLMAbunSD", "studyID","site.name","siteId","long", "lat", "reg","natInv","MaxYr", "MinYr", "dur", "AveSiteSST", "MinSiteSST", "MaxSiteSST")
for (i in 1:dim(sname)[1]) {
  dat<-subset(CRAB,sitename.studyname==rownames(sname)[i])
  m1<-glm(data=dat, RoundStandAbun~year, family="poisson") 
  m2<-glm(data=dat, StandAbunDecimal~year, family="quasipoisson") 
  m3<-glm.nb(data=dat, StandAbunDecimal~year) # negative binomial
  m4<-gls(data=dat, StandAbunDecimal~year) 
  mat[i,1:2]<-c(summary(m1)$coefficients[2],summary(m1)$coefficients[4])
  mat[i,3:4]<-c(summary(m2)$coefficients[2],summary(m2)$coefficients[4])
  mat[i,5:6]<-c(summary(m3)$coefficients[2],summary(m3)$coefficients[4])
  mat[i,7:8]<-c(summary(m4)$coefficients[2],as.data.frame(summary(m4)$tTable)[2,2])
  mat[i,9:12]<- c(dat$newID[1], rownames(sname)[i], dat$unique.site.id[1],dat$longitude[1])
  mat[i,13:17]<- c(dat$latitude[1], dat$region[1],dat$nat.invas[1],max(dat$year),min(dat$year))
  mat[i,18:21]<- c(length(unique(dat$year)),dat$AveSiteSST[1],dat$MinSiteSST[1],dat$MaxSiteSST[1]) }

```

Temperature models 

```{r eval=FALSE}
mat2<-as.data.frame(matrix(0,dim(sname)[1],3))
colnames(mat2)<- c("site.name","HADMaxSlope5", "HADMaxsd5") 
for (i in 1:dim(sname)[1]) {
  dat<-subset(CRAB,sitename.studyname==rownames(sname)[i])
  m1<-gls(data=dat, HADiSST.Max_1~year)
  mat2[i,1]<- rownames(sname)[i]
  mat2[i,2:3]<-c(summary(m1)$coefficients[2],as.data.frame(summary(m1)$tTable)[2,2])}

mat3<-data.frame(0,dim(sname)[1],3)
colnames(mat3)<- c("site.name","HADMinSlope5", "HADMinsd5") 
for (i in 1:dim(sname)[1]) {
  dat<-subset(CRAB,sitename.studyname==rownames(sname)[i])
  m1<-gls(data=dat, HADiSST.Min_1~year)
  mat3[i,1]<- rownames(sname)[i]  
  mat3[i,2:3]<-c(summary(m1)$coefficients[2],as.data.frame(summary(m1)$tTable)[2,2])}

mat4<-data.frame(0,dim(sname)[1],3)
colnames(mat4)<- c("site.name","AirMinSlope5", "AirMinsd5") 
for (i in 1:dim(sname)[1]) {
  dat<-subset(CRAB,sitename.studyname==rownames(sname)[i])
  m1<-gls(data=dat, AirT.Min~year)
  mat4[i,1]<- rownames(sname)[i]
  mat4[i,2:3]<-c(summary(m1)$coefficients[2],as.data.frame(summary(m1)$tTable)[2,2])}

mat5<-data.frame(0,dim(sname)[1],3)
colnames(mat5)<- c("site.name","AirMaxSlope5", "AirMaxsd") 
for (i in 1:dim(sname)[1]) {
  dat<-subset(CRAB,sitename.studyname==rownames(sname)[i])
  m1<-gls(data=dat, AirT.Max~year)
  mat5[i,1]<- rownames(sname)[i]
  mat5[i,2:3]<-c(summary(m1)$coefficients[2],as.data.frame(summary(m1)$tTable)[2,2])}

```

Put the dataframes together

```{r eval=FALSE}
slopes1 <- merge(mat,mat2, by="site.name")
slopes2 <- merge(slopes1,mat3, by="site.name")
slopes3 <- merge(slopes2,mat4, by="site.name")
slopes <- merge(slopes3,mat5, by="site.name")

```

# 2. Next, use slopes data from step 1 to build new models and choose best family 


```{r eval=FALSE}
atlantic<-filter(slopes,freg=="Atlantic")
native<-filter(slopes,freg=="Native")
atl.nat<-rbind(native,atlantic)
atl.nat$new.dist.code<-relevel(atl.nat$new.dist.code,ref="mid")
atl.nat$finv<-relevel(atl.nat$finv,ref="Native")

m1<-lme(data=atl.nat, PoisAbunSlope ~ scale(HADMaxSlope) + scale(HADMaxSlope):finv + scale(HADMaxSlope):new.dist.code, random=~1|fstudy, weights= varFixed(~PoisAbunSD))

m2<-lme(data=atl.nat, QuasAbunSlope ~ scale(HADMaxSlope) + scale(HADMaxSlope):finv + scale(HADMaxSlope):new.dist.code, random=~1|fstudy, weights= varFixed(~QuasAbunSD))

m3<-lme(data=atl.nat, NBAbunSlope ~ scale(HADMaxSlope) + scale(HADMaxSlope):finv + scale(HADMaxSlope):new.dist.code, random=~1|fstudy, weights= varFixed(~NBAbunSD))

m4<-lme(data=atl.nat, GLMAbunSlope ~ scale(HADMaxSlope) + scale(HADMaxSlope):finv + scale(HADMaxSlope):new.dist.code, random=~1|fstudy, weights= varFixed(~GLMAbunSD))

qqnorm(m1)
summary(m1)
## 1 and 3 have best qqnorm
## values in summary table of 1 and 3 very simolar
## 3 has slightly lower AIC

```

#2b. Examine all four temp variables with chosen family 

```{r}
m5<-lme(data=atl.nat, NBAbunSlope ~ scale(HADMaxSlope) + scale(HADMaxSlope):finv + scale(HADMaxSlope):new.dist.code, random=~1|fstudy, weights= varFixed(~NBAbunSD))

m6<-lme(data=atl.nat, NBAbunSlope ~ scale(HADMinSlope) + scale(HADMinSlope):finv + scale(HADMinSlope):new.dist.code, random=~1|fstudy, weights= varFixed(~NBAbunSD))

m7<-lme(data=atl.nat, NBAbunSlope ~ scale(AirMaxSlope) + scale(AirMaxSlope):finv + scale(AirMaxSlope):new.dist.code, random=~1|fstudy, weights= varFixed(~NBAbunSD))

m8<-lme(data=atl.nat, NBAbunSlope ~ scale(AirMinSlope) + scale(AirMinSlope):finv + scale(AirMinSlope):new.dist.code, random=~1|fstudy, weights= varFixed(~NBAbunSD))

qqnorm(m8)
summary(m8)
```


# 3. Next, use residual data from step 1 to build new models

```{r}
sname=data.frame(unique(CRAB$sitename.studyname));rownames(sname)<-sname[,1]
mat<-as.data.frame(matrix(0,dim(sname)[1],10000))

getabundanceTrends<-list()

for (i in 1:dim(sname)[1]) {
  i=1
  dat<-subset(CRAB,sitename.studyname==rownames(sname)[i])
  m1<-glm.nb(data=dat, StandAbunDecimal~year) # negative binomial
  getabundanceTrends[[i]] <- residuals(m1)
  }

getabundanceTrendsOUT<- do.call("rbind.fill", getabundanceTrends)
```


