---
title: "Modelling biotic and abiotic through time"
author: "BSB"
date: '2020-04-27'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("/Users/brandybiggar/Desktop/")
CRAB<-read.csv("egc_working_data_28.csv")
```


I have added a new study ID variable to standardize time and reduce the variance from duration. 

I then added a new ID variable wich is the newStudyID_sitename to ensure that the same site in different sections of the new study ID variable are treated as unique (not the same study). 

Now I have to check if there are still my required min of data for each studyID_sitename as a site of 4 years of data may have gotten split between 2 (or more) time sections leaving it with less than 4 years. 

-- There are 135 <4 years, so I will go remove those from the data. 

```{r}
library(dplyr)
a<- CRAB %>%
  group_by(sitename.studyname) %>%
  summarize(min=min(year), max=max(year), range=(max(year) - min(year) +1))
a
```



# Model

Use a loop to model every site.


```{r, ECHO=FALSE}
sname=data.frame(unique(CRAB$sitename.studyname));rownames(sname)<-sname[,1]
mat<-as.data.frame(matrix(0,dim(sname)[1],15))

for (i in 1:dim(sname)[1]) {
  dat<-subset(CRAB,sitename.studyname==rownames(sname)[i])
  m1<-glm(data=dat, RoundStandAbun~year, family="poisson") 
  mat[i,1]<- dat$newID[1]
  mat[i,2]<- rownames(sname)[i]
  mat[i,3]<- dat$unique.site.id[1]
  mat[i,4]<- dat$longitude[1]
  mat[i,5]<- dat$latitude[1]
  mat[i,6]<- dat$region[1]
  mat[i,7]<- dat$nat.invas[1]
  mat[i,8]<- max(dat$year)
  mat[i,9]<- min(dat$year)
  mat[i,10]<- length(unique(dat$year))
  mat[i,11]<- dat$AveSiteSST[1]
  mat[i,12]<- dat$MinSiteSST[1]
  mat[i,13]<- dat$MaxSiteSST[1]
  mat[i,14:15]<-c(summary(m1)$coefficients[2],as.data.frame(summary(m1)$tTable)[2,2]) 
}
which(is.na(mat))
```

# Clean up 

Label columns, add CIs.

```{r}
crabslopes<-as.data.frame(mat)
colnames(crabslopes)<- c("studyID","site.name","unique.site.id","longitude", "latitude", "region","nat.invas","Max_yr", "Min_yr","duration","AveSiteSST","MinSiteSST","MaxSiteSST", "AbunSlope", "AbunSD")
```


# Explore


```{r}
library(ggplot2)
hist(crabslopes$AbunSlope)
hist(crabslopes$AbunSD)
plot(crabslopes$AbunSD, crabslopes$AbunSlope) 
plot(crabslopes$duration, crabslopes$AbunSlope) 

```



Now repeat modelling loop with temperature ~ time


```{r}
sname=data.frame(unique(CRAB$sitename.studyname));rownames(sname)<-sname[,1]
mat<-data.frame(0,dim(sname)[1],3)

for (i in 1:dim(sname)[1]) {
  dat<-subset(CRAB,sitename.studyname==rownames(sname)[i])
  m1<-gls(data=dat, AirT.Max~year)
  mat[i,1]<-rownames(sname)[i]
  mat[i,2:3]<-c(summary(m1)$coefficients[2],as.data.frame(summary(m1)$tTable)[2,2]) # estimate, sd
}

temps<-as.data.frame(mat) %>%
colnames(temps)<- c("site.name","HADMinSlope", "HADMinsd") ## HAD Min

temps2<-as.data.frame(mat)
colnames(temps2)<- c("site.name","HADMaxSlope", "HADMaxsd") ## HAD Max

temps3<-as.data.frame(mat)
colnames(temps3)<- c("site.name","AirMinSlope", "AirMinsd") ## Air Min

temps4<-as.data.frame(mat)
colnames(temps4)<- c("site.name", "AirMaxSlope", "AirMaxsd") ## Air Max

tempslopes <- merge(temps,temps2, by="site.name")
tempslopes2 <- merge(tempslopes,temps3, by="site.name")
tempslopes3 <- merge(tempslopes2,temps4, by="site.name")

slopes.df<-merge(crabslopes,tempslopes3, by="site.name")

setwd("/Users/brandybiggar/Desktop/")
write.csv(slopes.df, "slope.data.May12.csv")
```

  m1<-glm(data=dat, HADiSST.Min_1~year)
  m2<-glm(data=dat, HADiSST.Max_1~year)
  m3<-glm(data=dat, AirT.Min~year)
  m4<-glm(data=dat, AirT.Max~year)
